---
title: "PCA"
output: html_notebook
---

# Principal Components Analyisi

Tutorials
[Data Camp - PCA Analysis in R](https://www.datacamp.com/tutorial/pca-analysis-r){.uri}

[FactoExtra](https://rpkgs.datanovia.com/factoextra/index.html){.uri} [STHDA: Principal Component Methods in R - Practical guide](http://www.sthda.com/english/articles/31-principal-component-methods-in-r-practical-guide/114-mca-multiple-correspondence-analysis-in-r-essentials/){.uri}

[Statistics globe: PCA before K Means Clustering](https://statisticsglobe.com/pca-before-k-means-clustering-r){.uri}\>

# Libraries

```{r}
library(tidyverse)
library(ggrepel) #load before factoextra
library(factoextra)
library(caret)
library(stats)
library(ggfortify)
library(ggplot2)
library(corrplot)
library(cowplot)
library(ggcorrplot)
library(ComplexHeatmap)
library(circlize)
library(ggExtra)
library("FactoMineR")
library(here)
```

# Process data

**Colors**

```{r}
condition_col = c(
                                "BBIBP (V3, D07)" = "#00A087FF",
                                "BBIBP (V3, D14)" = "#00A087FF",
                                "BBIBP (V3, D28)" = "#00A087FF",
                                "BNT (V1, D6)" = "#4DBBD5FF",
                                "BNT (V2, D1)" = "#4DBBD5FF",
                                "BNT (V3, D1)" = "#4DBBD5FF",
                                "BNT-I (D1)"= "#DC0000FF",
                                "BNT-I (D10, mild)"= "#DC0000FF",
                                "BNT-I (D10, severe)"= "#DC0000FF",
                                "BNT-I (D2)"= "#DC0000FF",
                                "BNT-I (D26, mild)"= "#DC0000FF",
                                "BNT-I (D26, severe)"= "#DC0000FF",
                                "BNT-I (D3)"= "#DC0000FF",
                                "BNT-I (D4)"= "#DC0000FF",
                                "BNT-I (D51, severe)"= "#DC0000FF",
                                "BNT-I-BNT (D51, mild)"= "#F5BB00",
                                "BNT-I-BNT (D51, severe)"= "#F5BB00",
                                "BNT-MO (V1, D6)"= "#72ddf7",
                                "BNT-MO (V3, D1)"= "#72ddf7",
                                "ChAd (V1, D3)"= "#3C5488FF",
                                "ChAd (V1, D6)"= "#3C5488FF",
                                "ChAd (V1, D7)"= "#3C5488FF",
                                "ChAd (V2, D1)"= "#3C5488FF",
                                "ChAd (V2, D3)"= "#3C5488FF",
                                "ChAd (V2, D7)"= "#3C5488FF",
                                "ChAd-BNT (V2, D0)"= "#4DBBD5FF",
                                "ChAd-BNT (V2, D3)"= "#4DBBD5FF",
                                "ChAd-BNT (V2, D7)"= "#4DBBD5FF",
                                "ChAd-BNT (V3, D1)"= "#4DBBD5FF",
                                "I (D1)"= "#DC0000FF",
                                "I (D10, moderate)"= "#DC0000FF",
                                "I (D10, severe)"= "#DC0000FF",
                                "I (D26, moderate)"= "#DC0000FF",
                                "I (D26, severe)"= "#DC0000FF",
                                "I (D51, moderate)"= "#DC0000FF",
                                "I (D51, severe)"= "#DC0000FF",
                                "I-BNT-I (D2)"= "#DC0000FF",
                                "I-BNT-I (D5)"= "#DC0000FF",
                                "I-I (D0)"= "#DC0000FF",
                                "I-I (D1)"= "#DC0000FF",
                                "I-I (D2)"= "#DC0000FF",
                                "I-I (D3)"= "#DC0000FF",
                                "I-I (D5)"= "#DC0000FF",
                                "ZF2001 (V3, D07)"= "#8491B4FF",
                                "ZF2001 (V3, D14)"= "#8491B4FF",
                                "ZF2001 (V3, D28)" = "#8491B4FF"
                              )
```

**Required files**

```{r Required files}
# Genes in long table format
all_degs_p_05_vac_infected_19_12_23 = read_csv(here("DGE analysis", "all_degs_p_05_vac_infected_19_12_23.csv"))

# Gene set annotation.Used only to merge and filter genes.
ImmuneGO_Annotated_Genes = readRDS(here("VaxGO gene sets", "ImmuneGO_Annotated_Genes_2024-03-26.rds"))

# Sample or condition annotation
ann_vaccines_19_1_24 <-
  read_csv(here("Annotations and metadata", "ann_vaccines_19-1-24.csv"))

data_annotation = ann_vaccines_19_1_24 %>%
  mutate(day = as.factor(day),
         dose = as.factor(dose)) 
```

**Filtering data**

```{r}
# Gene set data.
ImmuneGO_Annotated_Genes_all = ImmuneGO_Annotated_Genes
ImmuneGO_Annotated_Genes = ImmuneGO_Annotated_Genes_all %>%
  filter(!process %in% c("TCR REPERTOIRE", "BCR REPERTOIRE"))

ImmuneGO_Genes = all_degs_p_05_vac_infected_19_12_23 %>%
  inner_join(ImmuneGO_Annotated_Genes %>% 
               select(genes, process) %>% 
               distinct() %>% 
               select(genes),
               by = "genes") %>% 
  distinct()
```

**Input**

```{r}
#INPUT
data_genes = ImmuneGO_Genes 
filename = "ImmuneGO_Genes"
```

```{r Create dataframe with annotations and matrix for PCA}
#Convert long to wide table format.
#Dataframe with sample annotation
ann_vaccines_pca_matrix = data_genes %>% 
  mutate(log2fold_change = as.numeric(log2fold_change)) %>% 
  select(condition, genes, log2fold_change) %>% 
  pivot_wider(., names_from = "condition", 
                     values_from = "log2fold_change") %>% 
  replace(is.na(.), 0) %>%
  column_to_rownames("genes") %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("condition") %>% 
  inner_join(data_annotation %>% 
               select(condition, disease_vac, vaccine, day, week, dose, infection, type), by = "condition") 

# Convert dataframe to matrix without annotation columns
ann_vaccines_pca_matrix_ready = ann_vaccines_pca_matrix %>% 
  column_to_rownames("condition") %>% 
  select(!disease_vac:type) %>% 
  as.matrix()

```

# PCA

```{r PCA}
# Delete columns with near 0 variance
nearZeroVarCols <- nearZeroVar(ann_vaccines_pca_matrix_ready, saveMetrics = TRUE)
ann_vaccines_pca_matrix_ready <- ann_vaccines_pca_matrix_ready[, !nearZeroVarCols$nzv]
pca_res <- prcomp(ann_vaccines_pca_matrix_ready, scale. = T)

# Correlation plot ----
corr_matrix = cor(ann_vaccines_pca_matrix_ready %>% scale()) 
var <- get_pca_var(pca_res)

# corrplot = ggcorrplot(corr_matrix, hc.order = TRUE) + 
#   theme(axis.text.x = element_text(angle = 90, size = 5), 
#         axis.text.y = element_text(size = 5))
# #Salvar
# ggsave(corrplot, file = paste0(filename, "_corrplot.png"), 
#        width = 20, #Grande 20, pequeno 10
#        height= 20) #Grande 20, pequeno 10
# print(corrplot)

#PCA ----
data.pca = prcomp(corr_matrix) #PCA
summary(corr_matrix) #Retornar PCs

#Scree plot ----
scree_plot = fviz_eig(data.pca, 
                      addlabels = TRUE,
                      ylim = c(0, 70)) +
  geom_col(color = "#00AFBB", fill = "#00AFBB") +
  theme_classic()

#Save
ggsave(scree_plot, file = here(paste0(filename, "_screeplot.png"), "Figures"), width = 10, height = 3) 
print(scree_plot)

```

```{r Loadings plot}
# Loadings plot ----
options(ggrepel.max.overlaps = Inf)

circle_contrib= fviz_pca_var(pca_res, col.var = "cos2",
                              gradient.cols = c("black", "#DC0000FF"),
                              select.var= list(cos2 = 20), 
                              repel = T, 
                              labelsize = 4,
                              col.circle = NA) +
  xlab("") + 
  ylab("") +
  theme_minimal() +
theme(panel.grid = element_blank(),  # Remover linhas de grade
        axis.text = element_blank(),   # Remover r처tulos de texto dos eixos
        axis.ticks = element_blank()) 

#Save
ggsave(circle_contrib, file = here(paste0(filename, "_circle_contrib_wide.png"), "Figures"), width = 10, height = 5)
```

```{r KNN classification}
#KNN classification -----
#Use the screeplot generated % of explained variance for each dimension
scree_plot #Dim 1 = 68.2%, Dim 2 = 15.9%


# PC1-PC2 ------
#Determine the number of clusters
pca_scores <- data.frame(pca_res$x[, 1:2])
fviz_nbclust(pca_scores,  
                     FUNcluster = kmeans,
                     method = "wss")

pcas = "PC1-PC2"

set.seed(666) # Set seed for randomization
cluster_model <- kmeans(pca_res$x[, 1:2], centers = 3)  #Set the number of clusters

ann_vaccines_pca_matrix$cluster <- as.factor(cluster_model$cluster)
```

```{r PC1-PC2 scatter plot}
# Condition with density plot ------
pca_plot_knn_cluster = autoplot(pca_res, 
        data = ann_vaccines_pca_matrix, 
        colour = 'cluster')  +
  stat_ellipse(aes(color = cluster, fill = cluster), 
               geom = "polygon", 
               alpha = 0.2, 
               linetype = 1,
               size = 0.3,
               type = "t") +
  labs(title=paste0(filename, "_bycondition")) + 
  xlab("PC1 (68.2%)") +
  ylab("PC2 (15.9%)") +
  
  scale_color_manual(values = c("1" = "#DC0000FF", 
                                "2" = "#4DBBD5FF", 
                                "3" = "#8491B4FF", 
                                condition_col)) +  # Ellipse line colors
  scale_fill_manual(values = c("1" = "#DC0000FF", 
                                "2" = "#4DBBD5FF", 
                                "3" = "#8491B4FF")) +  # Cluster fill colors
  theme_minimal() +
  guides(col="none") +
  theme(panel.grid = element_blank(),
        text = element_text(color = "black")) +
  geom_text_repel(aes(label = condition,
                      segment.colour="gray70"),
                  box.padding = 0.3,
                  size = 3) 


ggsave(pca_plot_knn_cluster, 
       file = here(paste0(filename, pcas, "_KNN_Clustered_labels.png"), "Figures"), width = 10, height = 5)

pca_plot_knn_cluster_marginal = ggMarginal(
  pca_plot_knn_cluster,
  type = "histogram",
  groupFill = T,
  groupColour = T,
  xparams = list(binwidth = 0.1, size = 0.1),
  yparams = list(binwidth = 0.1, size = 0.1)
)

ggsave(
  pca_plot_knn_cluster_marginal,
  file = here(paste0(filename, pcas, "_KNN_Clustered_labels_Marginal.png"), "Figures"),
  width = 8,
  height = 5
)


```


# Heatmap

## Genes annotated

```{r}
####### INPUT
PC1 = var$cos2 %>%
  as.data.frame() %>%
  arrange(desc(Dim.1)) %>%
  filter(Dim.1 >= 0.5) %>% 
  #slice_head(n=20) %>% 
  rownames_to_column("genes") %>% 
  distinct() %>% 
  select(genes, cos2 = Dim.1) %>% 
    mutate(dimension = "PC1")

PC2 = var$cos2 %>%
  as.data.frame() %>%
  filter(Dim.2 >= 0.5) %>% 
  #slice_head(n=20) %>% 
  rownames_to_column("genes") %>% 
  distinct() %>% 
  select(genes, cos2 = Dim.2) %>% 
  mutate(dimension = "PC2")

PC1_2 = bind_rows(PC1, PC2) 
PC1_2 %>% 
  select(genes) %>% distinct()

PC1_2 %>% write.csv(file = "PC_1_2_cos2.csv", row.names = F)


# Gene annotation
ImmuneGO_Annotated_Genes
OtherGOs_Annotated_Genes

# INPUT

ann_genes = ImmuneGO_Annotated_Genes
pc = "PC2"
data_2 =  PC2 %>% 
  slice_max(order_by = cos2, n = 20) %>% 
  inner_join(data_genes, by = "genes")
filename_2 = paste0(filename, "_PCA_",pc, "_COVID")

######## Matrix
matrix = data_2 %>% 
  select(genes, condition, log2fold_change) %>%
  pivot_wider(names_from = "condition", 
              values_from = "log2fold_change", 
              values_fn = mean) %>% 
  replace(is.na(.), 0) %>%
  arrange(genes) %>% 
  column_to_rownames(var="genes") %>% 
  as.matrix()

######## Annotations
ann_vaccines_covid = data_2 %>% 
  select(condition) %>% 
  distinct() %>% 
  inner_join(ann_vaccines, by = "condition") %>% 
  select(condition, disease_vac, type, day) %>% 
  distinct()

#Columns
ann_cols_heatmap = matrix %>% 
  colnames() %>% 
  data.frame() %>% 
  filter(. != "genes") %>% 
  rename(condition = '.') %>% 
  inner_join(ann_vaccines_covid, by = "condition") %>% 
  distinct() %>% 
  arrange(condition) %>% 
  column_to_rownames("condition")

matrix_data = matrix %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column(var= "condition") %>% 
  inner_join(ann_cols_heatmap %>% rownames_to_column("condition"), by = "condition") %>% 
  select(!c(disease_vac:day)) %>% 
  arrange(condition) %>% 
  column_to_rownames("condition") %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate_if(is.character, as.numeric) %>%
  as.matrix()

#Rows Immune -----
ann_rows_immune = matrix_data %>%
  as.data.frame() %>%
  rownames_to_column("genes") %>%
  inner_join(ann_genes, by = "genes") %>%
  filter(go_term == "Manual") %>%
  select(genes, process) %>%
  distinct() %>%
  group_by(genes) %>%
  arrange(genes, factor(process, levels = c("INNATE IMMUNE SYSTEM", "ADAPTIVE IMMUNE SYSTEM", "OTHER", "ANTIMICROBIAL IMMUNE RESPONSE")), .by_group = TRUE) %>%
  mutate(i_process = row_number()) %>%
  pivot_wider(., names_from = "i_process", values_from = "process") %>%
  column_to_rownames("genes") %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("index") %>%
  arrange(desc(as.numeric(index))) %>%
  column_to_rownames("index") %>%
  t() %>%
  as.data.frame() %>%
  replace(is.na(.), ".")

# Rows Not immune -----
# ann_rows_notimmune = matrix_data %>%
#   as.data.frame() %>%
#   rownames_to_column("genes") %>%
#   left_join(ann_genes %>% 
#                select(process = term, genes = symbol, annotation) %>% 
#                filter(annotation == "Major process", 
#                       process != "Immune system"), by = "genes") %>% 
#   select(genes, process) %>%
#   distinct() %>%
#   group_by(genes) %>%
#   #arrange(genes, factor(process, levels = c("INNATE IMMUNE SYSTEM", "ADAPTIVE IMMUNE SYSTEM", "OTHER", "ANTIMICROBIAL IMMUNE RESPONSE")), .by_group = TRUE) %>% 
#   mutate(i_process = row_number()) %>% 
#   pivot_wider(., names_from = "i_process", values_from = "process") %>% 
#   column_to_rownames("genes") %>% 
#   t() %>% 
#   as.data.frame() %>% 
#   rownames_to_column("index") %>% 
#   arrange(desc(index)) %>% 
#   column_to_rownames("index") %>% 
#   t() %>% 
#   as.data.frame() %>% 
#   replace(is.na(.), ".")

#Verificar dimens천es do dataset
dim(matrix_data)
dim(ann_rows_immune)
dim(ann_cols_heatmap)
```

### Colors

### Plot

```{r}
#Heatmap annotation -----
ha = HeatmapAnnotation(df = ann_cols_heatmap, 
                       col = ann_vaxsig_colors,
                       annotation_name_side = "right")

row_ha = HeatmapAnnotation(df = ann_rows_immune, 
                            which = "row",
                       col = col_process_immune, #col_process_immune / col_process_notimmune 
                       show_annotation_name = F,
                       show_legend = F)

# Values colors -----
col_fun <- colorRamp2(c(-2, 0, 2), c("#9bc1bc", "white", "#ed6a5a"))


# Parameters -----
file = filename_2
mat = matrix_data
width_image = 30
height_image = 30
width = unit(15, "cm")
height = unit(10, "cm")
column_names_gp = gpar(fontsize = 10)
row_names_gp = gpar(fontsize = 10)
rect_gp = gpar(col = "gray80", lwd = 0.5)

fileimageheatmap_grouped = paste0(file, sep ="_", "Complexheatmap.png")

png(file = fileimageheatmap_grouped, 
    width = width_image, 
    height = height_image, 
    units="cm", 
    res=900)

heatmap_plot = Heatmap(mat,
                        top_annotation = ha,
                        left_annotation = row_ha,
                        show_row_names = TRUE,
                        row_names_side = "right",
                        show_heatmap_legend = TRUE,
                        heatmap_legend_param = list(direction = "horizontal"),
                        col = col_fun, #color
                        show_column_names = TRUE,
                        column_title = fileimageheatmap_grouped,
                        column_title_gp = gpar(fontsize = 10, fontface = "bold"),
                        column_names_rot = 45,
                        column_names_gp = column_names_gp,
                        rect_gp = rect_gp,
                        na_col = "white",
                        cluster_rows = TRUE,
                        row_names_gp = row_names_gp,
                        row_dend_width = unit(2, "cm"),
                        column_split = 3,
                        row_split = 2,
                        row_gap = unit(2, "mm"),
                        cluster_columns = TRUE,
                        column_gap = unit(8, "mm"),
                        show_column_dend = F,
                        show_row_dend = F,
                        row_dend_side = "left",
                        column_dend_side = "top",
                        clustering_distance_rows = "euclidean",
                        border = TRUE,
                        width = width,
                        height = height 
)

draw(heatmap_plot, 
     annotation_legend_side = "left", 
     heatmap_legend_side = "left")
dev.off()

```

### STATS

```{r}
stats_1 = covid_other_immunego_genes %>% 
  inner_join(ann_vaccines_covid_other, by = "condition") %>% 
  group_by(genes, covid_other) %>% 
  summarise(n_conditions = n()) %>% 
  filter(all(c("COVID", "OTHER") %in% covid_other)) %>% 
  arrange(desc(n_conditions), genes) %>% 
  ungroup()


ggplot_stats1 = ggplot(stats_1) +
  aes(x = reorder(genes, n_conditions), fill = covid_other, weight = n_conditions) +
  geom_bar() +
  theme_minimal() +
  scale_fill_manual(
    values = c(COVID = "#56cfe1",
    OTHER = "#343a40")
  ) +
  theme_minimal() +
  labs(fill = "Covid or non-covid") +
  xlab("Genes") + 
  ylab("Number of shared conditions") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1, size = 8),
        axis.text.y = element_text(size = 7, angle = 0),
        panel.grid.minor.y = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size = 12, face = "bold")) +
  # geom_text(aes(label = n_conditions, y = n_conditions), 
  #           position = position_dodge(width = 0.9), 
  #           vjust = 0.2, 
  #           hjust = -0.2, 
  #           size = 1,
  #           angle = 90) +
  ggtitle("Genes shared") +
  coord_flip()

ggsave(ggplot_stats1, file = "Covid_Vax_genes_shared_column.png", width = 10, height = 20)




stats_2 = matrix_data %>%
  as.data.frame() %>%
  rownames_to_column("genes") %>%
  inner_join(ImmuneGO_Annotated_Genes, by = "genes") %>%
  filter(go_term == "Manual",
         !(process %in% c("ADAPTIVE IMMUNE SYSTEM",
                          "INNATE IMMUNE SYSTEM"))) %>%
  select(genes, process, immune_system) %>%
  distinct() %>%
  group_by(genes, immune_system) %>% 
  summarise(n_processes = n()) %>% 
  arrange(desc(n_processes)) %>% 
  ungroup() %>% 
  filter(n_processes >1)
  
ggplot_stats2 = ggplot(stats_2) +
  aes(x = reorder(genes, n_processes), fill = immune_system, weight = n_processes) +
  geom_bar() +
  theme_minimal() +
  scale_fill_manual(
    values = c(Adaptive = "#6f2dbd",
    Innate = "#06d6a0",
    Complement = "#ffadc7")
  ) +
  theme_minimal() +
  labs(fill = "Immune system") +
  xlab("Genes") + 
  ylab("Number of shared conditions") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1, size = 8),
        axis.text.y = element_text(size = 7, angle = 0),
        panel.grid.minor.y = element_blank(),
        strip.text.x.top = element_text(size = 10, face = "bold"),
        legend.text = element_text(size=12),
        legend.title = element_text(size = 12, face = "bold")) +
  # geom_text(aes(label = n_conditions, y = n_conditions), 
  #           position = position_dodge(width = 0.9), 
  #           vjust = 0.2, 
  #           hjust = -0.2, 
  #           size = 1,
  #           angle = 90) +
  ggtitle("Genes shared") +
  coord_flip() +
  facet_wrap(~immune_system, scales = "free_y")

ggsave(ggplot_stats2, file = "ImmuneGO_genes_shared_major_process.png", width = 6, height = 20)

```

# Principal Variance Components Analysis

## PVCA of normalized raw counts

```{r}
#### PVCA of normalized raw counts -----
# BiocManager::install("ExpressionNormalizationWorkflow")
library(ExpressionNormalizationWorkflow)

#Covariants
ann_vaccines_samples = ann_vaccines_samples_4_12_23
cvrts = ann_vaccines_samples %>% 
  arrange(condition) %>%
  distinct() %>% 
  column_to_rownames("sample")

#Expression data

exprs = all_matrices_normalized_counts %>% 
  replace(is.na(.), 0) %>% 
  replace(., . == -Inf, 0) %>% 
  replace(., . == Inf, 0) %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("sample") %>% 
  arrange(sample) %>% 
  inner_join(cvrts %>% 
               #filter(disease_vac == "V") %>% 
               rownames_to_column("sample"),
               select(sample) %>% 
               distinct(), 
             by = "sample") %>% 
  column_to_rownames("sample") %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate_if(is.character, as.numeric) %>% 
  replace(., . == -Inf, 0) %>% 
  replace(., . == Inf, 0) %>% 
  replace(is.na(.), 0)

cvrts = cvrts %>% 
  rownames_to_column("sample") %>% 
  inner_join(exprs %>% 
               colnames() %>% 
               as.data.frame() %>% 
               select(sample = "."), 
             by = "sample") %>% 
  arrange(sample) %>% 
  column_to_rownames("sample")

# Input
cvrts_eff_var = c("day","age", "sex", "dose", "type", "regimen") #For vaccination only (no ongoing infection)
cvrts_eff_var = c("day","age", "sex", "dose", "type", "previous_vaccination","variant", "prior_infection", "severity", "disease_vac", "regimen") #For vaccination and infection
inpData <- expSetobj(exprs, cvrts) #Convert to Expression Dataset Object
pct_thrsh <- 0.75  #Threshold

#PVCA analysis
pvca = pvcAnaly(inpData, pct_thrsh, cvrts_eff_var) 

pvca_df = pvca$dat %>% 
  as.data.frame() %>% 
  t() %>%
  as.data.frame() %>% 
  rename(wavgpv = V1) %>% 
  cbind(pvca$label %>% as.data.frame() %>% rename(batch = '.')) %>% 
  as.data.frame() %>% 
  mutate(batch = as.factor(batch),
         wavgpv = as.numeric(wavgpv),
         batch = fct_rev(fct_reorder(batch, wavgpv))) %>% 
  arrange(batch, wavgpv)

# Visualize
pvca_df_plot = ggplot(pvca_df) +
  aes(x = batch, y = wavgpv) +
  geom_col(fill = "#5e60ce") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust=1, color = "black"),
        axis.text.y = element_text(size = 20, vjust = 1, hjust=1, color = "black")) +
  geom_text(aes(label = round(wavgpv,3), y = wavgpv), 
            position = position_dodge(width = 0.9), 
            vjust = -0.2, 
            hjust = 0.5, 
            size = 4,
            angle = 0) 

pvca_df_plot

ggsave(pvca_df_plot, file = "pvca_raw_normalized_immunegenes_allconditions.png", width = 12, height = 10)

```

## PVCA of log2FC

```{r}
#### PVCA of log2FC -----

# BiocManager::install("ExpressionNormalizationWorkflow")

library(ExpressionNormalizationWorkflow)

#Covariants
cvrts = ann_vaccines %>% 
  select(gse_id, condition:previous_vaccination) %>% 
  arrange(condition) %>%
  filter(day != 0) %>% 
  distinct() %>% 
  column_to_rownames("condition")
#Expression data

degs_lg2fc_wide = degs_updown_filtered_p_05_vac_infected %>% 
  select(condition, genes, log2fold_change) %>%
  distinct() %>% 
  arrange(condition) %>% 
  pivot_wider(names_from = "condition", values_from = "log2fold_change", values_fn = mean)
  
exprs = degs_lg2fc_wide %>% 
  column_to_rownames("genes") %>% 
  replace(is.na(.), 0) %>% 
  replace(., . == -Inf, 0) %>% 
  replace(., . == Inf, 0) %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("condition") %>% 
  arrange(condition) %>% 
  inner_join(cvrts %>% 
               rownames_to_column("condition"),
               select(condition) %>% 
               distinct(), 
             by = "condition") %>% 
  column_to_rownames("condition") %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate_if(is.character, as.numeric) %>% 
  replace(., . == -Inf, 0) %>% 
  replace(., . == Inf, 0) %>% 
  replace(is.na(.), 0)

cvrts = cvrts %>% 
  rownames_to_column("condition") %>% 
  inner_join(exprs %>% 
               colnames() %>% 
               as.data.frame() %>% 
               select(condition = "."), 
             by = "condition") %>% 
  column_to_rownames("condition")

# Input
cvrts_eff_var = c("day", "dose", "prior_infection", "previous_vaccination", "severity", "disease_vac", "regimen", "type" )
inpData <- expSetobj(exprs, cvrts)
pct_thrsh <- 0.75 
pvca = pvcAnaly(inpData, pct_thrsh, cvrts_eff_var)

pvca_df = pvca$dat %>% 
  as.data.frame() %>% 
  t() %>%
  as.data.frame() %>% 
  rename(wavgpv = V1) %>% 
  cbind(pvca$label %>% as.data.frame() %>% rename(batch = '.')) %>% 
  as.data.frame() %>% 
  mutate(batch = as.factor(batch),
         wavgpv = as.numeric(wavgpv),
         batch = fct_rev(fct_reorder(batch, wavgpv))) %>% 
  arrange(batch, wavgpv)

# Visualize

pvca_df_plot = ggplot(pvca_df) +
  aes(x = batch, y = wavgpv) +
  geom_col(fill = "#5e60ce") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 20, angle = 45, vjust = 1, hjust=1, color = "black"),
        axis.title.x = element_text(size = 20,  color = "black"))

ggsave(pvca_df_plot, file = "pvca_log2foldchange.png", width = 12, height = 10)


```


# PCA: COVID-19 and others, only up and down regulated genes


```{r Required files}

# Gene set annotation.Used only to merge and filter genes.
ImmuneGO_Annotated_Genes
ImmuneGO_Genes = degs_covid_vax %>% 
  select(-direction)


# Sample or condition annotation
data_annotation = VAXSigDB_Annotated %>% 
  clean_names() %>% 
  select(condition = vaccine_condition, type = platform, microbe_type, target_pathogen_disease, time) %>% 
  mutate(time = if_else(time == "ANY", "0", time),
         time = if_else(time == "1-3", "2", time),
         time = as.numeric(time)) %>% 
  bind_rows(ann_vaccines_covid_other %>% rename(time = day_total)) %>% 
  select(condition, type, target_pathogen_disease, time)

VAXSigDB_Annotated_days = VAXSigDB_Annotated %>% 
  select(VACCINE_CONDITION, `DATE-TIME`:TIME) %>% 
  mutate(TIME = case_when(TIME == "ANY" ~ "0",
                          TIME == "1-3" ~ "2", 
                          TRUE ~ TIME),
         TIME = as.numeric(TIME)) %>% 
 mutate(day = case_when(DATE == "H" ~ TIME/24,
                        DATE == "M" ~ TIME * 30,
                        DATE == "W" ~ TIME * 7,
                        TRUE ~ TIME)) %>% 
  select(condition = VACCINE_CONDITION, day) %>% 
  distinct()

sars = data_annotation %>% 
  filter(target_pathogen_disease == "SARS") %>% 
  mutate(target_pathogen_disease = "COVID-19") %>% 
  rename(day = time)

data_annotation = data_annotation %>% 
  left_join(VAXSigDB_Annotated_days, by = "condition") %>% 
  filter(day < 100,
         day > 0) %>% 
  distinct() %>% 
  select(-time) %>% 
  bind_rows(sars) %>% 
  filter(type != "I") %>% 
  mutate(target_pathogen_disease = if_else(target_pathogen_disease == "COVID-19", "COVID-19", "OTHER"))

```

## PCA

```{r}
#INPUT
data_genes = ImmuneGO_Genes 
filename = "ImmuneGO_Genes_UP_DOWN_COVID_OTHER"
```

```{r Create dataframe with annotations and matrix for PCA}
#Convert long to wide table format.
#Dataframe with sample annotation
ann_vaccines_pca_matrix = data_genes %>% 
  mutate(regulation_numeric = as.numeric(regulation_numeric)) %>% 
  select(genes, condition, regulation_numeric) %>% 
  distinct() %>%
  pivot_wider(names_from = "condition", 
              values_from = "regulation_numeric",
              values_fn = mean) %>% 
  replace(is.na(.), 0) %>%
  column_to_rownames("genes") %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column("condition") %>% 
  inner_join(data_annotation, by = "condition") %>% 
  distinct() 

# Convert dataframe to matrix without annotation columns
ann_vaccines_pca_matrix_ready = ann_vaccines_pca_matrix %>% 
  column_to_rownames("condition") %>% 
  select(!c(type, target_pathogen_disease, day)) %>% 
  as.matrix()

```

```{r PCA}
# Delete columns with near 0 variance
nearZeroVarCols <- nearZeroVar(ann_vaccines_pca_matrix_ready, saveMetrics = TRUE)
ann_vaccines_pca_matrix_ready <- ann_vaccines_pca_matrix_ready[, !nearZeroVarCols$nzv]
pca_res <- prcomp(ann_vaccines_pca_matrix_ready)

# Correlation plot ----
corr_matrix = cor(ann_vaccines_pca_matrix_ready) 
var <- get_pca_var(pca_res)

#PCA ----
#data.pca = prcomp(corr_matrix) #PCA
summary(corr_matrix) #Retornar PCs

#Scree plot ----
scree_plot = fviz_eig(data.pca, 
                      addlabels = TRUE,
                      ylim = c(0, 80)) +
  geom_col(color = "#00AFBB", fill = "#00AFBB") +
  theme_classic()

#Save
ggsave(scree_plot, file = paste0(filename, "_screeplot.png"), width = 10, height = 3) 
print(scree_plot)

```

```{r Loadings plot}
# Loadings plot ----
options(ggrepel.max.overlaps = Inf)

# Filter top vars for each PC
cos2_df <- get_pca_var(pca_res)$cos2
cos2_df <- as.data.frame(cos2_df)
cos2_df$variable <- rownames(cos2_df)
top_n_cos2 <- function(df, n = 10) {
  df %>%
    arrange(desc(cos2)) %>%
    slice_head(n = n)
}
top_cos2_Dim1 <- top_n_cos2(cos2_df %>% select(variable, cos2 = Dim.1))
top_cos2_Dim2 <- top_n_cos2(cos2_df %>% select(variable, cos2 = Dim.2))
top_cos2_combined <- bind_rows(top_cos2_Dim1, top_cos2_Dim2) %>%
  distinct(variable, .keep_all = TRUE)



circle_contrib= fviz_pca_var(pca_res, col.var = "cos2",
                              gradient.cols = c("black", "#DC0000FF"),
                              select.var = list(name = top_cos2_combined$variable),
                              repel = T, 
                              labelsize = 4,
                              col.circle = NA) +
  xlab("") + 
  ylab("") +
  theme_minimal() +
theme(panel.grid = element_blank(),  # Remover linhas de grade
        axis.text = element_blank(),   # Remover r처tulos de texto dos eixos
        axis.ticks = element_blank()) 

circle_contrib


#Save
ggsave(circle_contrib, file = paste0(filename, "_circle_contrib_wide.png"), width = 10, height = 5)

```

```{r KNN classification}
#KNN classification -----
#Use the screeplot generated % of explained variance for each dimension
scree_plot #Dim 1 = 58.8%, Dim 2 = 10.5%


# PC1-PC2 ------
#Determine the number of clusters
pca_scores <- data.frame(pca_res$x[, 1:2])
fviz_nbclust(pca_scores,  
                     FUNcluster = kmeans,
                     method = "wss")

pcas = "PC1-PC2"

set.seed(666) # Set seed for randomization
cluster_model <- kmeans(pca_res$x[, 1:2], centers = 2)  #Set the number of clusters

ann_vaccines_pca_matrix$cluster <- as.factor(cluster_model$cluster)
```

```{r}
# Condition with density plot ------
pca_plot_knn_cluster = autoplot(pca_res, 
        data = ann_vaccines_pca_matrix, 
        size = 2,
        shape = 'type',
        color = "day")  +
  stat_ellipse(aes(fill = cluster), 
               geom = "path", 
               alpha = 0.1, 
               linetype = 1,
               size = 1,
               type = "t") +
  labs(title=paste0(filename, "_bycondition")) + 
  xlab("PC1 (58.8%)") +
  ylab("PC2 (10.5%)") +
  scale_color_stepsn(colors = c("#DC0000FF", "#F39B7FFF", "#4DBBD5FF"),
                     limits = c(0, 14), 
                     na.value = "#4DBBD5FF") +
  scale_shape_manual(values = c("VLP" = 1,
           "LA" =  5,
           "CONJ" = 3,
           'IN'= 4, #1st Gen  vaccines
           'VV' = 2, #3rd Gen vaccines
           'RNA' = 6,
           'SU' = 7,
           'I'= 8)) +
  theme_minimal() +
  theme(panel.grid = element_blank(),
        text = element_text(color = "black")) 

pca_plot_knn_cluster 

ggsave(pca_plot_knn_cluster, file = paste0(filename, pcas, "_KNN_Clustered_labels.png"), width = 10, height = 5)
```

```{r}
pca_plot_knn_cluster_interactive = autoplot(pca_res, 
        data = ann_vaccines_pca_matrix, 
        shape = 'type',
        color = "day")  +
  labs(title=paste0(filename, "_bycondition")) + 
  xlab("PC1 (58.8%)") +
  ylab("PC2 (10.5%)") +
  scale_color_gradientn(colors = c("darkblue", "purple"), limits = c(0, 14), na.value = "purple") +
  scale_shape_manual(values = c("VLP" = 1,
                                 "LA" =  2,
                                 "CONJ" = 3, 
                                 'IN'= 4, #1st Gen  vaccines
                                 'VV' = 5, #3rd Gen vaccines
                                 'RNA' = 6,
                                 'SU' = 7)) +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(color = "black")) 

pca_plot_knn_cluster_interactive %>% ggplotly()
```



## Heatmap

### Genes annotated

```{r}
####### INPUT
PC1 = var$cos2 %>%
  as.data.frame() %>%
  arrange(desc(Dim.1)) %>%
  #slice_head(n=20) %>% 
  rownames_to_column("genes") %>% 
  distinct() %>% 
  select(genes, cos2 = Dim.1) %>% 
    mutate(dimension = "PC1")

PC2 = var$cos2 %>%
  as.data.frame() %>%
  filter(Dim.2 >= 0.5) %>% 
  #slice_head(n=20) %>% 
  rownames_to_column("genes") %>% 
  distinct() %>% 
  select(genes, cos2 = Dim.2) %>% 
  mutate(dimension = "PC2")

PC1_2 = bind_rows(PC1, PC2) 
PC1_2 %>% 
  select(genes) %>% distinct()

PC1_2 %>% write.csv(file = "PC_1_2_cos2.csv", row.names = F)


# Gene annotation
ImmuneGO_Annotated_Genes
OtherGOs_Annotated_Genes

# INPUT

ann_genes = ImmuneGO_Annotated_Genes
pc = "PC2"
data_2 =  PC2 %>% 
  slice_max(order_by = cos2, n = 20) %>% 
  inner_join(data_genes, by = "genes")
filename_2 = paste0(filename, "_PCA_",pc, "_COVID")

######## Matrix
matrix = data_2 %>% 
  select(genes, condition, log2fold_change) %>%
  pivot_wider(names_from = "condition", 
              values_from = "log2fold_change", 
              values_fn = mean) %>% 
  replace(is.na(.), 0) %>%
  arrange(genes) %>% 
  column_to_rownames(var="genes") %>% 
  as.matrix()

######## Annotations
ann_vaccines_covid = data_2 %>% 
  select(condition) %>% 
  distinct() %>% 
  inner_join(ann_vaccines, by = "condition") %>% 
  select(condition, disease_vac, type, day) %>% 
  distinct()

#Columns
ann_cols_heatmap = matrix %>% 
  colnames() %>% 
  data.frame() %>% 
  filter(. != "genes") %>% 
  rename(condition = '.') %>% 
  inner_join(ann_vaccines_covid, by = "condition") %>% 
  distinct() %>% 
  arrange(condition) %>% 
  column_to_rownames("condition")

matrix_data = matrix %>% 
  t() %>% 
  as.data.frame() %>% 
  rownames_to_column(var= "condition") %>% 
  inner_join(ann_cols_heatmap %>% rownames_to_column("condition"), by = "condition") %>% 
  select(!c(disease_vac:day)) %>% 
  arrange(condition) %>% 
  column_to_rownames("condition") %>% 
  t() %>% 
  as.data.frame() %>% 
  mutate_if(is.character, as.numeric) %>%
  as.matrix()

#Rows Immune -----
ann_rows_immune = matrix_data %>%
  as.data.frame() %>%
  rownames_to_column("genes") %>%
  inner_join(ann_genes, by = "genes") %>%
  filter(go_term == "Manual") %>%
  select(genes, process) %>%
  distinct() %>%
  group_by(genes) %>%
  arrange(genes, factor(process, levels = c("INNATE IMMUNE SYSTEM", "ADAPTIVE IMMUNE SYSTEM", "OTHER", "ANTIMICROBIAL IMMUNE RESPONSE")), .by_group = TRUE) %>%
  mutate(i_process = row_number()) %>%
  pivot_wider(., names_from = "i_process", values_from = "process") %>%
  column_to_rownames("genes") %>%
  t() %>%
  as.data.frame() %>%
  rownames_to_column("index") %>%
  arrange(desc(as.numeric(index))) %>%
  column_to_rownames("index") %>%
  t() %>%
  as.data.frame() %>%
  replace(is.na(.), ".")

# Rows Not immune -----
# ann_rows_notimmune = matrix_data %>%
#   as.data.frame() %>%
#   rownames_to_column("genes") %>%
#   left_join(ann_genes %>% 
#                select(process = term, genes = symbol, annotation) %>% 
#                filter(annotation == "Major process", 
#                       process != "Immune system"), by = "genes") %>% 
#   select(genes, process) %>%
#   distinct() %>%
#   group_by(genes) %>%
#   #arrange(genes, factor(process, levels = c("INNATE IMMUNE SYSTEM", "ADAPTIVE IMMUNE SYSTEM", "OTHER", "ANTIMICROBIAL IMMUNE RESPONSE")), .by_group = TRUE) %>% 
#   mutate(i_process = row_number()) %>% 
#   pivot_wider(., names_from = "i_process", values_from = "process") %>% 
#   column_to_rownames("genes") %>% 
#   t() %>% 
#   as.data.frame() %>% 
#   rownames_to_column("index") %>% 
#   arrange(desc(index)) %>% 
#   column_to_rownames("index") %>% 
#   t() %>% 
#   as.data.frame() %>% 
#   replace(is.na(.), ".")

#Verificar dimens천es do dataset
dim(matrix_data)
dim(ann_rows_immune)
dim(ann_cols_heatmap)
```

### Plot

```{r}
#Heatmap annotation -----
ha = HeatmapAnnotation(df = ann_cols_heatmap, 
                       col = ann_vaxsig_colors,
                       annotation_name_side = "right")

row_ha = HeatmapAnnotation(df = ann_rows_immune, 
                            which = "row",
                       col = col_process_immune, #col_process_immune / col_process_notimmune 
                       show_annotation_name = F,
                       show_legend = F)

# Values colors -----
col_fun <- colorRamp2(c(-2, 0, 2), c("#9bc1bc", "white", "#ed6a5a"))


# Parameters -----
file = filename_2
mat = matrix_data
width_image = 30
height_image = 30
width = unit(15, "cm")
height = unit(10, "cm")
column_names_gp = gpar(fontsize = 10)
row_names_gp = gpar(fontsize = 10)
rect_gp = gpar(col = "gray80", lwd = 0.5)

fileimageheatmap_grouped = paste0(file, sep ="_", "Complexheatmap.png")

png(file = fileimageheatmap_grouped, 
    width = width_image, 
    height = height_image, 
    units="cm", 
    res=900)

heatmap_plot = Heatmap(mat,
                        top_annotation = ha,
                        left_annotation = row_ha,
                        show_row_names = TRUE,
                        row_names_side = "right",
                        show_heatmap_legend = TRUE,
                        heatmap_legend_param = list(direction = "horizontal"),
                        col = col_fun, #color
                        show_column_names = TRUE,
                        column_title = fileimageheatmap_grouped,
                        column_title_gp = gpar(fontsize = 10, fontface = "bold"),
                        column_names_rot = 45,
                        column_names_gp = column_names_gp,
                        rect_gp = rect_gp,
                        na_col = "white",
                        cluster_rows = TRUE,
                        row_names_gp = row_names_gp,
                        row_dend_width = unit(2, "cm"),
                        column_split = 3,
                        row_split = 2,
                        row_gap = unit(2, "mm"),
                        cluster_columns = TRUE,
                        column_gap = unit(8, "mm"),
                        show_column_dend = F,
                        show_row_dend = F,
                        row_dend_side = "left",
                        column_dend_side = "top",
                        clustering_distance_rows = "euclidean",
                        border = TRUE,
                        width = width,
                        height = height 
)

draw(heatmap_plot, 
     annotation_legend_side = "left", 
     heatmap_legend_side = "left")
dev.off()

```

### STATS

```{r}
stats_1 = covid_other_immunego_genes %>% 
  inner_join(ann_vaccines_covid_other, by = "condition") %>% 
  group_by(genes, covid_other) %>% 
  summarise(n_conditions = n()) %>% 
  filter(all(c("COVID", "OTHER") %in% covid_other)) %>% 
  arrange(desc(n_conditions), genes) %>% 
  ungroup()


ggplot_stats1 = ggplot(stats_1) +
  aes(x = reorder(genes, n_conditions), fill = covid_other, weight = n_conditions) +
  geom_bar() +
  theme_minimal() +
  scale_fill_manual(
    values = c(COVID = "#56cfe1",
    OTHER = "#343a40")
  ) +
  theme_minimal() +
  labs(fill = "Covid or non-covid") +
  xlab("Genes") + 
  ylab("Number of shared conditions") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1, size = 8),
        axis.text.y = element_text(size = 7, angle = 0),
        panel.grid.minor.y = element_blank(),
        legend.text = element_text(size=12),
        legend.title = element_text(size = 12, face = "bold")) +
  # geom_text(aes(label = n_conditions, y = n_conditions), 
  #           position = position_dodge(width = 0.9), 
  #           vjust = 0.2, 
  #           hjust = -0.2, 
  #           size = 1,
  #           angle = 90) +
  ggtitle("Genes shared") +
  coord_flip()

ggsave(ggplot_stats1, file = "Covid_Vax_genes_shared_column.png", width = 10, height = 20)




stats_2 = matrix_data %>%
  as.data.frame() %>%
  rownames_to_column("genes") %>%
  inner_join(ImmuneGO_Annotated_Genes, by = "genes") %>%
  filter(go_term == "Manual",
         !(process %in% c("ADAPTIVE IMMUNE SYSTEM",
                          "INNATE IMMUNE SYSTEM"))) %>%
  select(genes, process, immune_system) %>%
  distinct() %>%
  group_by(genes, immune_system) %>% 
  summarise(n_processes = n()) %>% 
  arrange(desc(n_processes)) %>% 
  ungroup() %>% 
  filter(n_processes >1)
  
ggplot_stats2 = ggplot(stats_2) +
  aes(x = reorder(genes, n_processes), fill = immune_system, weight = n_processes) +
  geom_bar() +
  theme_minimal() +
  scale_fill_manual(
    values = c(Adaptive = "#6f2dbd",
    Innate = "#06d6a0",
    Complement = "#ffadc7")
  ) +
  theme_minimal() +
  labs(fill = "Immune system") +
  xlab("Genes") + 
  ylab("Number of shared conditions") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, vjust = 1.5, hjust=1, size = 8),
        axis.text.y = element_text(size = 7, angle = 0),
        panel.grid.minor.y = element_blank(),
        strip.text.x.top = element_text(size = 10, face = "bold"),
        legend.text = element_text(size=12),
        legend.title = element_text(size = 12, face = "bold")) +
  # geom_text(aes(label = n_conditions, y = n_conditions), 
  #           position = position_dodge(width = 0.9), 
  #           vjust = 0.2, 
  #           hjust = -0.2, 
  #           size = 1,
  #           angle = 90) +
  ggtitle("Genes shared") +
  coord_flip() +
  facet_wrap(~immune_system, scales = "free_y")

ggsave(ggplot_stats2, file = "ImmuneGO_genes_shared_major_process.png", width = 6, height = 20)

``` 